IMPC Mouse data - Variance in sex differences

### 2020-04-20 Prep & old/SydneySexDiff_Felix_11-2-2020TEST.Rmd
Add files via upload

```{r}
library(readr)
library(dplyr)
library(metafor)
library(devtools)
library(purrr)
library(tidyverse)
library(tibble)
library(kableExtra)
library(robumeta)
library(ggpubr)
library(ggplot2)
library(here)
```

Functions for preparing the data for meta analyses

1) "Population statistics": "calculate_population_stats"
This function groups animals from the same strain and same insitiution together. This is done for each trait seoarately, and only for traits that have been measured in both sexes. Any group containing fewer than 5 individuals is excluded.

@@

2) Extraction of effect sizes and sample variances: "create_meta_analysis_effect_sizes"

@@

3) Calculate meta-analysis statistics

Based on function created by A M Senior @ the University of Otago NZ 03/01/2014:

@@

1) Data loading and cleaning of the csv file

@@

    # subset to reasonable set of variables
    # date_of_experiment: Jeremy suggested using as an indicator of batch-level effects
    select(production_center, strain_name, strain_accession_id, biological_sample_id, pipeline_stable_id, procedure_group, procedure_name, sex, date_of_experiment, age_in_days, weight, parameter_name, data_point) %>%

@@

    # filter to groups with > 1 centre  
    # **Review** - retaining merge here to retain ordering of data used in previous version
    # usually i would use mutate instead of creating sperate dataset
    merge(tmp, 
          tmp %>% group_by(parameter_name) %>%
    summarise(center_per_trait = length(unique(production_center, na.rm = TRUE)))
        )%>%
    filter(center_per_trait >= 2) %>%

@@

    # add grouping variable: these were decided based on functional groups and procedures 
    mutate(parameter_group = group$parameter[match(parameter_name, group$parameter_name)] ) %>%
    # usually i would use left_join
        # left_join(by="parameter_name", 
    #   read_csv(here("data", "ParameterGrouping.csv")) %>%
    #     select(-id)
    # ) %>%

@@

Check length of different variables / sample sizes:

```{r}
length(unique(data$parameter_name)) # 232 traits
length(unique(data$parameter_group)) # 161 parameter groups
length(unique(data$procedure_name)) # 26 procedure groups
length(unique(data$biological_sample_id)) # 27147 individial mice   #FZ added 11-12-2019
#number of males and females per strain per production center #FZ added 11-12-2019  # FOR TABLE SAMPLE SIZES
data %>% group_by(production_center, strain_name) %>% count(biological_sample_id, sex) %>% count(sex) %>% print(n = Inf)

# Phase 1: across all traits

Create function for sub-setting the data to choose only one data point per individual per trait. This is a necessary step for the loop across all traits

data_subset_parameterid_individual_by_age <- function(mydata, parameter, age_min=0, age_center=100) {
  tmp <- mydata %>%
    filter(
      age_in_days >= age_min,
      id == parameter
    ) %>%
    # take results for single individual closest to age_center
    mutate(age_diff = abs(age_center - age_in_days)) %>%
    group_by(biological_sample_id) %>%
    filter(age_diff == min(age_diff)) %>%
    select(-age_diff)# %>% 
#    filter(!duplicated(biological_sample_id))

Loop running meta-analysis on all traits

@@

(n <- length(unique(data$id)))

@@

--------
**REVIEW**: The above loop gives some errors, can these be fixed?
**FZ**: In the above function, we use 'tryCatch' and 'conditionMessage' to prevent the loop from aborting when the first error at row 84 is produced.
**FZ**  Convergence in the two listed non-converging cases can't be achieved by sensibly tweaking (other optim etc.).
**FZ**  Since we only learn about non-convergence in the loop, it's difficult to exclude the two traits beforehand.
**FZ**  Similarly, the 8 traits with very low variation (160, ..., 231) are hard to exclude before.
**FZ**  Warnings are ok: indicating cases where variance components are set to zero during likelihood optimization.

ERROR : 84 Optimizer (optim) did not achieve convergence (convergence = 10). 
ERROR : 158 Optimizer (optim) did not achieve convergence (convergence = 10). 
ERROR : 160 NA/NaN/Inf in 'y' 
ERROR : 161 NA/NaN/Inf in 'y' 
ERROR : 162 NA/NaN/Inf in 'y' 
ERROR : 163 NA/NaN/Inf in 'y' 
ERROR : 165 NA/NaN/Inf in 'y' 
ERROR : 166 NA/NaN/Inf in 'y' 
ERROR : 168 NA/NaN/Inf in 'y'
ERROR : 231 NA/NaN/Inf in 'y'

Also a bunch of warnings, are any of these a concern?

> warnings()
Warning messages:
1: In metafor::rma.mv(yi = effect_size_CVR, V = sample_variance_CVR,  ... :
  Rows with NAs omitted from model fitting.
...
13: In metafor::rma.mv(yi = effect_size_RR, V = sample_variance_RR,  ... :
  There are outcomes with non-positive sampling variances.
14: In metafor::rma.mv(yi = effect_size_RR, V = sample_variance_RR,  ... :
  'V' appears to be not positive definite.
16: In metafor::rma.mv(yi = effect_size_CVR, V = sample_variance_CVR,  ... :
  Single-level factor(s) found in 'random' argument. Corresponding 'sigma2' value(s) fixed to 0.
...
50: In metafor::rma.mv(yi = effect_size_RR, V = sample_variance_RR,  ... :
  'V' appears to be not positive definite

------
Now that we have a "results" table with each of the meta-analytic means for all effect sizes of interest, we can use this table as part of the Shiny App, which will then be able to back calculate the percentage differences between males and females for mean, variance and coefficient of variance. We'll export and use this in the Shiny App. **Note that I have not dealt with convergence issues in some of these models, and so, this will need to be done down the road**

(Note Susi 31/7/2019: This dataset contains dublicated values, plus no info on what the "traits" mean. I will change Dan N's to one further belwo, that have been cleaned up already
FILE TO USE: METACOMBO (around line 500))

@@

Procedure names, grouping variables etc. are merged back together with the results from the metafor analysis above. This requires loading of another excel sheet, "procedures.csv"

```{r}
results_alltraits_grouping2 <- 
  results_alltraits_grouping %>% 
  left_join(by="id",
             data %>% select(id, parameter_group, procedure = procedure_name, procedure_name, parameter_name) %>% #susi: check parateer group
              # REVIEW -- this bext line replciates precious code, but removes some procedures, is that right?
		  # FZ: No. This creates the data from the complete data set ('data'), which was not carried over into 'results_alltraits_grouping'
		  # FZ: We filter duplicated id's to get only one unique row per id (and there is one id per parameter_name)
              filter(!duplicated(id))
            ) %>%
  # FZ: Below we add 'procedure' (from the previously loaded 'procedures.csv') as a variable
  left_join(by="procedure", 
            procedures %>% distinct()
            )

n <- length(unique(results_alltraits_grouping2$parameter_name)) # 232
```

Removal of traits that did not achieve convergence, are nonsensical for analysis of variance (such as traits that show variation, such as number of ribs, digits, etc). 14 traits from the originally 232 that had been included are removed.

**Review**: How did we determine this the ones with problems? 
**FZ**: In addition to the sentence one line above, to do: add information about cause of exclusion ( 1. non-convergence of metafor models, or 2. no variation/nonsensical); reference traits to be excluded by trait name (not id number)
**FZ**: excluded because of non-convergence: 84, 158
**FZ**: excluded because no variation/nonsensical: 144, 158, 160,... 
```{r}

#FZ delete# has_problems <- c(84, 144, 158, 160, 161, 162, 163, 165, 166, 167, 168, 221, 222, 231)
#FZ delete# meta_clean <- results_alltraits_grouping2 %>% filter(!id %in% has_problems)
#FZ delete# meta_problem <- results_alltraits_grouping2 %>% filter(id %in% has_problems)

#**FZ**: We exclude 14 parameter names for which metafor models didn't converge ("dp t cells", "mzb (cd21/35 high)"), and of parameters that don't harbour enough variation
meta_clean <- results_alltraits_grouping2 %>% 
	  filter(!parameter_name %in% c("dp t cells", "mzb (cd21/35 high)", "number of caudal vertebrae", 
	  "number of cervical vertebrae", "number of digits", "number of lumbar vertebrae", "number of pelvic vertebrae", "number of ribs left",                       
        "number of ribs right", "number of signals", "number of thoracic vertebrae", "total number of acquired events in panel a",
        "total number of acquired events in panel b", "whole arena permanence"))

```
```

**Reveiw**: check against old script -- identical, remove once fixed

```{r, eval=FALSE}
meta_clean.test <- readRDS("../meta_clean.test.rds")  #FZ comment: doesn't work for me. '../' is for the github location?
all.equal(meta_clean, meta_clean.test)
all.equal(meta_clean %>% select(-parameter_group), meta_clean.test %>% mutate(id=as.integer(id), GroupingTerm = as.character(GroupingTerm)))
```

## Meta-analysis, Phase 2: non-independent traits
### Dealing with Correlated Parameters, preparation

This dataset contained a number of highly correlated traits, such as different kinds of cell counts (for example, hierarchical parameterization within immunological assays). As those data-points are not independent of each other,  we conducted a meta analyses on these correlated parameters to collapse the number of levels.

@@

meta1 <- meta_clean
length(unique(meta1$procedure)) #18
length(unique(meta1$GroupingTerm)) #9
# Review: next line now 155 not 148 #FZ comment: still 148 for me.
length(unique(meta1$parameter_group)) # 148 levels. To be used as grouping factor for meta-meta analysis / collapsing down based on things that are classified identically in "parameter_group" but have different "parameter_name"
length(unique(meta1$parameter_name)) #218
```

@@

This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same  parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation.

@@

kable(cbind(meta1 %>% count(parameter_group))) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```

```{r}
meta1_sub <- meta1 %>%
  # add summary of number of parameter names in each parameter group
  group_by(parameter_group) %>%
  mutate(par_group_size = length(unique(parameter_name)), 
         sampleSize = as.numeric(sampleSize)) %>% 
  ungroup() %>% 
  # Create subsets with > 1 count (par_group_size > 1)
  filter(par_group_size > 1) # 90 observations
```

#### Meta-analyses on correlated (sub-)traits, using robumeta` 
The subset of the data is prepared (nested), and in this first step the model of the meta analysis effect sizes are calculated

meta1b <-
  meta1 %>%
  group_by(parameter_group) %>% 
  summarize(par_group_size = length(unique(parameter_name, na.rm = TRUE)))
#this gives a summary of number of parameter names in each parameter group, now it neeeds to get merged it back together

@@

# nesting
n_count <- meta1_sub %>%
  group_by(parameter_group) %>%
  mutate(raw_N = sum(sampleSize)) %>%
  nest() %>%
  ungroup()

@@

Extract and save parameter estimates:
Function to collect the outcomes of the "mini" meta analysis

@@

Extraction of values created during Meta analysis using robu meta:

@@

```{r}
kable(cbind(robu_all, robu_all)) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```

Merge the two data sets (the new [robu_all] and the initial [uncorrelated sub-traits with count = 1])

```{r}
meta_all <- meta1 %>%
  filter(par_group_size == 1) %>%
  as_tibble()
# str(meta_all)
# str(robu_all)
# which(is.na(match(names(meta_all),names(robu_all))))  # check
```

#### Combine data

```{r}
# Step1 (columns are matched by name (in our case, 'parameter_group'), and any missing columns will be filled with NA) #FZ added explanation of bibd_rows
combinedmeta <- bind_rows(robu_all, meta_all)
# glimpse(combinedmeta)

# Steps 2&3 (add information about number of traits in a parameter group, procedure, and grouping term) #FZ added expalnation in brackets
metacombo <- combinedmeta
metacombo$counts <- meta1$par_group_size[match(metacombo$parameter_group, meta1$parameter_group)] 
metacombo$procedure2 <- meta1$procedure[match(metacombo$parameter_group, meta1$parameter_group)]
metacombo$GroupingTerm2 <- meta1$GroupingTerm[match(metacombo$parameter_group, meta1$parameter_group)]

kable(cbind(metacombo, metacombo)) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```

@@

```{r}
metacombo <- metacombo[, c(1, 21:23, 2:13)] # SUSI CHANGE TO NAMES!!!
names(metacombo)[3] <- "procedure" #change "3" to name
names(metacombo)[4] <- "GroupingTerm" #change "4" to name

# Quick pre-check before doing plots

metacombo %>%
  group_by(GroupingTerm) %>%
  dplyr::summarize(MeanCVR = mean(lnCVR), MeanVR = mean(lnVR), MeanRR = mean(lnRR))

```{r}
# SHINY APP # Now that we have a corrected "results" table with each of the meta-analytic means for all effect sizes of interest, we can use this table as part of the Shiny App, which will then be able to back calculate the percentage differences between males and females for mean, variance and coefficient of variance. We'll export and use this in the Shiny App. **Note that I have not dealt with convergence issues in some of these models, and so, this will need to be done down the road**

## Note Susi 31/7/2019: This has been cleaned up already
# FILE TO USE: METACOMBO
### note: to use

# trait_meta_results <- write.csv(metacombo, file = "export/trait_meta_results.csv")
```

## Meta-analysis, Phase 3: Second-order meta analysis for functional groups
#### Perform meta-analyses (3 for each of the 9 grouping terms: lnCVR, lnVR, lnRR) 
This is the full result dataset

kable(cbind(metacombo, metacombo)) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```

#### Prepare data
Nesting, calculating the number of parameters within each grouping term, and running the meta-analysis

```{r}
metacombo_final <- metacombo %>%
  group_by(GroupingTerm) %>%
#FZ comment: changed 'nest' to 'nest_legacy' to keep old syntax/functionality
  nest_legacy()

@@

# For all grouping terms
metacombo_final_all <- metacombo %>%
#FZ comment: changed 'nest' to 'nest_legacy' to keep old syntax/functionality
  nest_legacy()

@@

Re-structure data for each grouping term; delete unused variables  #FZ comment: referencing of cells below doesn't depend on previous oreding of the data, i.e. only changes if output structure from metafor::rma.uni changes

```{r}
Behaviour <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Behaviour") %>% mutate(
  lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
  lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
  lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
))[, c(1, 7:18)]

Immunology <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Immunology") %>% mutate(
  lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
  lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
  lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
))[, c(1, 7:18)]

Hematology <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Hematology") %>% mutate(
  lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
  lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
  lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
))[, c(1, 7:18)]

Hearing <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Hearing") %>% mutate(
  lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
  lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
  lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
))[, c(1, 7:18)]

Physiology <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Physiology") %>% mutate(
  lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
  lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
  lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
))[, c(1, 7:18)]

Metabolism <- as.data.frame(overall1 %>% filter(., GroupingTerm == "Metabolism") %>% mutate(
  lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
  lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
  lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
))[, c(1, 7:18)]

>>>>

```{r}
library(readr)
library(dplyr)
library(metafor)
library(devtools)
library(purrr)
library(tidyverse)
library(tidyr)
library(tibble)
library(kableExtra)
library(robumeta)
library(ggpubr)
library(ggplot2)
library(here)
```

## Functions 
for preparing the data for meta analyses

### 1) Subsetting data
Create function for sub-setting the data to choose only one data point per individual per trait: "data_subset_parameterid_individual_by_age"

```{r}
data_subset_parameterid_individual_by_age <- function(mydata, parameter, age_min=0, age_center=100) {
  tmp <- mydata %>%
    filter(
      age_in_days >= age_min,
      id == parameter
    ) %>%
    # take results for single individual closest to age_center
    mutate(age_diff = abs(age_center - age_in_days)) %>%
    group_by(biological_sample_id) %>%
    filter(age_diff == min(age_diff)) %>%
    select(-age_diff)# %>% 
#    filter(!duplicated(biological_sample_id))  #Felix 6/2/2020: this line can be deleted

### 2) "Population statistics"
Create function called: "calculate_population_stats"
This function groups animals from the same strain and same insitiution together. This is done for each trait seoarately, and only for traits that have been measured in both sexes. Any group containing fewer than 5 individuals is excluded.

@@

### 3) Extraction of effect sizes and sample variances
Function: "create_meta_analysis_effect_sizes"

@@

### 4) Calculate meta-analysis statistics

Based on a function created by A M Senior @ the University of Otago NZ 03/01/2014:

@@

### 1) Data loading and cleaning of the csv file

@@

    # subset to reasonable set of variables, date_of_experiment used as an indicator of batch-level effects
    select(production_center, strain_name, strain_accession_id, biological_sample_id, pipeline_stable_id, procedure_group, procedure_name, sex, date_of_experiment, age_in_days, weight, parameter_name, data_point) %>%

@@

    # filter to groups with > 1 centre  
    merge(tmp, 
          tmp %>% group_by(parameter_name) %>%
    summarise(center_per_trait = length(unique(production_center, na.rm = TRUE)))
        )%>%
    filter(center_per_trait >= 2) %>%

@@

    # add grouping variable: these were decided based on functional groups and procedures 
    mutate(parameter_group = group$parameter[match(parameter_name, group$parameter_name)] ) %>%

@@

Checking length of different variables and sample sizes.

### Table 1:  "Strains and Center Sample Sizes"
This table summarises the available numbers of male and female mice from each strain and originating institution.

```{r echo = FALSE, results = 'hold'}
#length(unique(data$parameter_name)) # 232 traits
#length(unique(data$parameter_group)) # 161 parameter groups
#length(unique(data$procedure_name)) # 26 procedure groups
#length(unique(data$biological_sample_id)) # 27147 individial mice

#number of males and females per strain per production center 
kable(cbind(data %>% group_by(production_center, strain_name) %>% count(biological_sample_id, sex) %>% count(sex) %>% print(n = Inf))) %>%
  kable_styling() %>%
  scroll_box(width = "70%", height = "200px")

## SZ: I don't understand why the tibble is showing in the knitted html
```

# Meta-analyses
## 1. Population as analysis unit 
(Step C, Figure 3 in main document)

### Loop: Meta-analyses on all traits

@@

n <- length(unique(data$id))

@@

In the above function, we use 'tryCatch' and 'conditionMessage' to prevent the loop from aborting when the first error at row 84 is produced.
As convergence in the two listed non-converging cases can't be achieved by sensibly tweaking (other optim etc.), and we only learn about non-convergence in the loop, it is not possible to exclude the traits (N=2) beforehand.
Similarly, there are 8 traits with very low variation, which can not be excluded prior to running the loop.

The produced "Warnings" indicate cases where variance components are set to zero during likelihood optimization.

@@

Procedure names, grouping variables and trait names ("parameter_names") are merged back together with the results from the metafor analysis above.

```{r}
results_alltraits_grouping2 <- 
  results_alltraits_grouping %>% 
  left_join(by="id",
             data %>% select(id, parameter_group, procedure = procedure_name, procedure_name, parameter_name) %>%   # We filter duplicated id's to get only one unique row per id (and there is one id per parameter_name)
              filter(!duplicated(id))
            ) %>%
  # Below we add 'procedure' (from the previously loaded 'procedures.csv') as a variable
  left_join(by="procedure", 
            procedures %>% distinct()
            )

#(n <- length(unique(results_alltraits_grouping2$parameter_name))) # 232
```

### Removal of traits

14 traits from the originally 232 that had been included are removed because they either did not achieve convergence or are nonsensical for analysis of variance (such as traits that show no variation, see list below).

Not converged: "dp t cells", "mzb (cd21/35 high)"

Not enough variation: "number of caudal vertebrae", "number of cervical vertebrae", "number of digits", "number of lumbar vertebrae", "number of pelvic vertebrae", "number of ribs left","number of ribs right", "number of signals", "number of thoracic vertebrae", "total number of acquired events in panel a","total number of acquired events in panel b", "whole arena permanence".

```{r}
# We exclude 14 parameter names for which metafor models didn't converge ("dp t cells", "mzb (cd21/35 high)"), and of parameters that don't harbour enough variation
meta_clean <- results_alltraits_grouping2 %>% 
	  filter(!parameter_name %in% c("dp t cells", "mzb (cd21/35 high)", "number of caudal vertebrae", 
	  "number of cervical vertebrae", "number of digits", "number of lumbar vertebrae", "number of pelvic vertebrae", "number of ribs left",                       
        "number of ribs right", "number of signals", "number of thoracic vertebrae", "total number of acquired events in panel a",
        "total number of acquired events in panel b", "whole arena permanence"))

## 2. Meta-analysis: condensing non-independent traits 
(Step F in Figure 3 in main article)

### Dealing with Correlated Parameters, preparation

This dataset contained a number of highly correlated traits, such as different kinds of cell counts (for example hierarchical parameterization within immunological assays). As those data-points are not independent of each other,  we conducted meta analyses on these correlated parameters to collapse the number of levels.

@@

meta1 <- meta_clean 
length(unique(meta1$procedure)) #18
length(unique(meta1$GroupingTerm)) #9
length(unique(meta1$parameter_group)) # 148 levels. To be used as grouping factor for meta-meta-analysis / collapsing down based on things that are classified identically in "parameter_group" but have different "parameter_name"
length(unique(meta1$parameter_name)) #218
```

@@

### Table 2: Numbers of correlated and uncorrelated traits
This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation.

```{r}
kable(cbind(meta1 %>% count(parameter_group))) %>%
  kable_styling() %>%
  scroll_box(width = "60%", height = "200px")
```

```{r}
meta1_sub <- meta1 %>%
  # Add summary of number of parameter names in each parameter group
  group_by(parameter_group) %>%
  mutate(par_group_size = length(unique(parameter_name)), 
         sampleSize = as.numeric(sampleSize)) %>% 
  ungroup() %>% 
  # Create subsets with > 1 count (par_group_size > 1)
  filter(par_group_size > 1) # 90 observations
```

#### Meta-analyses on correlated (sub-)traits, using robumeta` 
Here we pepare the subset of the data (using nest()), and in this first step the model of the meta-analysis effect sizes are calculated

@@

# Create summary of number of parameter names in each parameter group, and merge back together

meta1b <-
  meta1 %>%
  group_by(parameter_group) %>% 
  summarize(par_group_size = length(unique(parameter_name, na.rm = TRUE)))

@@

# Nesting and meta-analyses on correlated traits, using robumeta

n_count <- meta1_sub %>%
  group_by(parameter_group) %>%
  mutate(raw_N = sum(sampleSize)) %>%
  nest() %>%
  ungroup()

@@

#### Extracting and save parameter estimates
Here we apply an additional Function to collect the outcomes of the 'mini-meta-analysis' that has ondensed our non-independent traits.

@@

Extraction of values created during meta-analyses using robumeta

@@

#### Combining data 
Merge the two data sets (the new [robu_all] and the initial [uncorrelated sub-traits with count = 1])

```{r}
meta_all <- meta1 %>%
  filter(par_group_size == 1) %>%
  as_tibble()
# glimpse(meta_all)
# glimpse(robu_all)

# Step 1:  Columns are matched by name (in our case, 'parameter_group'), and any missing columns will be filled with NA
combinedmeta <- bind_rows(robu_all, meta_all)
# glimpse(combinedmeta)

# Steps 2&3: Add information about number of traits in a parameter group, procedure, and grouping term
metacombo <- combinedmeta
metacombo$counts <- meta1$par_group_size[match(metacombo$parameter_group, meta1$parameter_group)] 
metacombo$procedure2 <- meta1$procedure[match(metacombo$parameter_group, meta1$parameter_group)]
metacombo$GroupingTerm2 <- meta1$GroupingTerm[match(metacombo$parameter_group, meta1$parameter_group)]

@@

```{r}
metacombo <- metacombo[c("parameter_group", "counts","procedure2","GroupingTerm2", "lnCVR","lnCVR_lower","lnCVR_upper","lnCVR_se","lnVR","lnVR_lower","lnVR_upper","lnVR_se","lnRR","lnRR_lower","lnRR_upper","lnRR_se")]

names(metacombo)[names(metacombo)=="procedure2"] <- "procedure" 
names(metacombo)[names(metacombo)=="GroupingTerm2"] <- "GroupingTerm"

# Quick pre-check before doing plots
metacombo %>%
  group_by(GroupingTerm) %>%
  dplyr::summarize(MeanCVR = mean(lnCVR), MeanVR = mean(lnVR), MeanRR = mean(lnRR))
```

### Table 3: Full corrected dataset

We use this corrected (for correlated traits) "results" table, which contains each of the meta-analytic means for all effect sizes of interest, for further analyses.  We further use this table as part of the Shiny App, which is able to provide the percentage differences between males and females for mean, variance and coefficient of variance.

This is the full result dataset
```{r}
kable(metacombo) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

# trait_meta_results <- write.csv(metacombo, file = "export/trait_meta_results.csv")  #Felix 7/2/2020: I think this can be deleted for publication!
```

## 3. Second-order meta-analysis for functional groups
(Section H in Figure 3 in main article)

### Performing meta-analyses (3 for each of the 9 grouping terms: lnCVR, lnVR, lnRR)

#### Preparation of data

Nesting, calculating the number of parameters within each grouping term, and running the meta-analyses

```{r}
metacombo_final <- metacombo %>%
  group_by(GroupingTerm) %>%
   nest()

@@

# For all grouping terms
metacombo_final_all <- metacombo %>%
  nest(data = everything())

@@

### Re-structuring the data for each grouping term
We here delete unused variables, and select the respective effect sizes. Please note - the referencing of the cells does NOT depend on previous ordering of the data. This would only be affected if the output structure from metafor::rma.uni changes.

```{r}
Behaviour <- overall1 %>% 
  filter(., GroupingTerm == "Behaviour") %>% 
  mutate(
    lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
    lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
    lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
  )  %>%
  select(., GroupingTerm, lnCVR:lnRR_se)

Immunology <- overall1 %>% 
  filter(., GroupingTerm == "Immunology") %>% 
  mutate(
    lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
    lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
    lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
  )  %>%
  select(., GroupingTerm, lnCVR:lnRR_se)

Hematology <- overall1 %>% 
  filter(., GroupingTerm == "Hematology") %>% 
  mutate(
    lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
    lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
    lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
  )  %>%
  select(., GroupingTerm, lnCVR:lnRR_se)

Hearing <- overall1 %>% 
  filter(., GroupingTerm == "Hearing") %>% 
  mutate(
    lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
    lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
    lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
  )  %>%
  select(., GroupingTerm, lnCVR:lnRR_se)

Physiology <- overall1 %>% 
  filter(., GroupingTerm == "Physiology") %>% 
  mutate(
    lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
    lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
    lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
  )  %>%
  select(., GroupingTerm, lnCVR:lnRR_se)

Metabolism <- overall1 %>% 
  filter(., GroupingTerm == "Metabolism") %>% 
  mutate(
    lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
    lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
    lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
  )  %>%
  select(., GroupingTerm, lnCVR:lnRR_se)

Morphology <- overall1 %>% 
  filter(., GroupingTerm == "Morphology") %>%
  mutate(
    lnCVR = .[[4]][[1]]$b, lnCVR_lower = .[[4]][[1]]$ci.lb, lnCVR_upper = .[[4]][[1]]$ci.ub, lnCVR_se = .[[4]][[1]]$se,
    lnVR = .[[5]][[1]]$b, lnVR_lower = .[[5]][[1]]$ci.lb, lnVR_upper = .[[5]][[1]]$ci.ub, lnVR_se = .[[5]][[1]]$se,
    lnRR = .[[6]][[1]]$b, lnRR_lower = .[[6]][[1]]$ci.lb, lnRR_upper = .[[6]][[1]]$ci.ub, lnRR_se = .[[6]][[1]]$se
  )  %>%
  select(., GroupingTerm, lnCVR:lnRR_se)
